---
title: "dplyr - Part II"
author: "Author"
date: "Date"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE) #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
library(tidyverse)
library(palmerpenguins)
data(package = "palmerpenguins")
library(knitr)
library(kableExtra)
```

## Outline

--

- Continue working with `dplyr`

--

- `filter()`: subset rows using column values

--

- `group_by()`: group by one or more variables

--

- `summarise()`: summarise each group to fewer rows

--

- Exercises

---

## Filter(): Subset rows using column values

--

- The following expressions are mostly used:

  - `==`, `>`, etc.

  - `&` *(and)*
  
  - `|` *(or)*
  
  - `!` *(is not)*
  
  - `xor()` *(elementwise exclusive OR)*
  
  - `is.na()`
  
  -  `between()` *(shortcut for `x >= left` & `x <= right`)*
  
---

#### Quantitative variables

```{r}
penguins %>% 
  filter(bill_length_mm > 40) %>% #<<
  head(n = 2) %>% 
  kable() %>%  kable_styling(font_size = 14)
```

---

#### Qualitative variables

```{r}
penguins %>% 
  filter(species == "Adelie") %>% #<<
  head(n = 3) %>% 
  kable() %>%  kable_styling(font_size = 14)
```

--

#### Boolean operators

```{r}
penguins %>% 
  filter(species == "Adelie" & island == "Dream") %>% #<< 
  head(n = 1) %>% 
  kable() %>%  kable_styling(font_size = 14)
```

---

```{r}
penguins %>% 
  filter(species == "Adelie" & year != 2008) %>% #<<
  head(n = 2) %>% 
  kable() %>%  kable_styling(font_size = 14)
```

--

- `xor()`

```{r}
penguins %>% 
  filter(xor(species == "Adelie", flipper_length_mm > 200)) %>% #<<
  head(n = 2) %>% 
  kable() %>%  kable_styling(font_size = 14)
```

---

- `is.na()`

```{r}
penguins %>% 
  filter(is.na(body_mass_g)) %>% #<<
  head(n = 2) %>% 
  kable() %>% kable_styling(font_size = 14)
```

--

- `between()`

```{r}
penguins %>% 
  filter(between(body_mass_g, 5500, 6000)) %>% #<<
  head(n = 2) %>% 
  kable() %>%  kable_styling(font_size = 14)
```


---

## Quick recap

--

- So far, we know how to...

--

- subset columns with `select()`

--

- modify existing and create new columns with `mutate()`

--

- rename columns with `rename()`

--

- change the order of columns  with `relocate()`

--

- subset rows with `filter()`

--

- **Questions?**

---

class: center, middle

<h3>
How do we work with different groups in our datasets?
</h3>

---

## Group_by: Group by one or more variables

--

- Examples of groups in the `Palmerpenguins` dataset:

--

```{r}
distinct(penguins, species)
```

--

```{r}
distinct(penguins, sex)
```

---

- Group_by is a powerful function allowing operations per group (which is usually more interesting)

--

- What is the average weigth of the penguins per year?

--

```{r}
penguins_weigth <- penguins %>% 
  group_by(year) %>% #<<
  mutate(weigth_avg = mean(body_mass_g, na.rm = TRUE))

distinct(penguins_weigth, weigth_avg) %>% 
  kable()
```

---

- What is the maximum bill length of the three different penguin species?

--

```{r}
penguins_bill <- penguins %>% 
  group_by(species) %>% #<<
  mutate(bill_length = max(bill_length_mm, na.rm = TRUE))

distinct(penguins_bill, bill_length) %>% 
  kable()
```

---

- You can also group by several groups

--

- What is the maximum bill length of the three different penguin species on the three different islands?

--

```{r}
penguins_bill_island <- penguins %>% 
  group_by(species, island) %>% #<<
  mutate(bill_length_max = max(bill_length_mm, na.rm = TRUE))

distinct(penguins_bill_island, bill_length_max) %>% 
  kable()
```

--

- There's a tidier way to obtain the same result

---

## Summarise(): Summarise each group to fewer rows

--

- Let's look at the same examples again

--

- What is the average weigth of the penguins per year?

--

```{r}
penguins %>% 
  group_by(year) %>%
  summarise(weigth_avg = mean(body_mass_g, na.rm = TRUE)) %>% #<<
  kable()
```

---

- What is the maximum bill length of the three different penguin species on the three different islands?

--

```{r}
penguins %>% 
  group_by(species, island) %>% #<<
  summarise(bill_length_max = max(bill_length_mm, na.rm = TRUE)) %>% #<<
  kable()
```

---

- Again, combining different operations is possible

--

```{r}
penguins %>% 
  group_by(species) %>% 
  summarise(bill_length_max = max(bill_length_mm, na.rm = TRUE), #<<
            bill_length_min = min(bill_length_mm, na.rm = TRUE), #<<
            average_weigth = mean(body_mass_g, na.rm = TRUE)) %>% #<<
  kable()
```

---

class: center, middle

<h3>
Questions?
</h3>

---

## Recap & outlook

--

- Introduced to `dplyr`

--

- `select()`: picks variables based on their names.

--

- `mutate()`: adds new variables that are functions of existing variables

--

- `filter()`: picks cases based on their values.

--

- `group_by`: allows operations across groups

--

- `summarise()`: reduces multiple values down to a single summary

--

- Next up: Data visualisation with `ggplot2` and applying everything in a final exercise

---

class: center, middle

<h3>
Time to exercise!
</h3>

```{r, include=FALSE}
pagedown::chrome_print("dplyr_part2.html")
```
