---
title: "dplyr (Part One) - Exercises"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE) #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
```

## Exercises

### `mpg`

- Load the `dplyr` package – either by loading the specific package or the `tidyverse` package as a whole.

- Make sure you have the most recent version of `dpylr` (1.0.2) to ensure smooth coding.
 
- Load the `ggplot2` package as well if you haven’t done so yet. Store the default dataset `mpg` (Fuel economy data from 1999 to 2008 for 38 popular models of cars) in a data frame.
 
- Change the name of the following variables: `displ` to displacement, `cyl` to cylinders, `trans` to transmission, `fl` to fueltype.

- Drop the column `class`.

- Create a new column `distance_diff`, which shows the difference between cty and hwy. What's the highest difference in absolute terms?

---

- Create another new column `fivegal_distance`, which indicates how far one could drive on the highway with 5 gallons in their tank. How many columns do you have in your dataset by now?

- Bonus question: Which model(s) would get you the furthest on the highway with one gallon in the tank? 

### `starwars`

- Load the starwars dataset (part of `dplyr`) and store it in a data frame.

- Drop all columns after `species`.

- Create a new column `height_m`, listing the character’s height in meters. Drop `height` afterwards.

- Keep the following variables: `name`, `mass`, all `_color` variables as well as all numeric variables.

- Set the height variable as the first column in your dataset. Put all the `_color`-variables at the end of the dataset. Try to find a neater way to do that than just manually typing all the `_color`-variables. Tip: Check out the `relocate()`-helpfile.

---

- Bonus question: Which characters were born more than 100 years before the Battle of Yavin?

```{r, include=FALSE}
pagedown::chrome_print("dplyr1_exercises.html")
```



