---
title: "dplyr (Part One) - Solutions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE, warning = FALSE, results = "hide") #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
```

## Exercises

### `mpg`

- Load the `dplyr` package – either by loading the specific package or the `tidyverse` package as a whole.

```{r}
library(tidyverse)
```

- Make sure you have the most recent version of `dpylr` (1.0.2) to ensure smooth coding.

```{r}
sessionInfo()
```
 
---

- Load the `ggplot2` package as well if you haven’t done so yet. Store the default dataset `mpg` (Fuel economy data from 1999 to 2008 for 38 popular models of cars) in a data frame.

```{r}
#library(ggplot2) #if dplyr loaded specifically
cars <- mpg
```

- Change the name of the following variables: `displ` to displacement, `cyl` to cylinders, `trans` to transmission, `fl` to fueltype.

```{r}
cars <- cars %>% 
  rename(displacement = displ, 
         cylinders = cyl, 
         transmission = trans, 
         fueltype = fl)
```

- Drop the column `class`.

```{r}
cars <- cars %>% 
  select(-class)
```

---

- Create a new column `distance_diff`, which shows the difference between cty and hwy. What's the highest difference in absolute terms?

```{r, results='markup'}
cars <- cars %>% 
  mutate(distance_diff = cty-hwy)

max(abs(cars$distance_diff))
```

- Create another new column `fivegal_distance`, which indicates how far one could drive on the highway with 5 gallons in their tank. How many columns do you have in your dataset by now?

```{r, results='markup'}
cars <- cars %>% 
  mutate(fivegal_distance = hwy*5)

ncol(cars)
```

---

- Bonus question: Which car model(s) would get you the furthest on the highway with one gallon in the tank? 

```{r, results='markup'}
cars %>% 
  filter(hwy == max(hwy)) %>% 
  pull(model)
```

---

### `starwars`

- Load the starwars dataset (part of `dplyr`) and store it in a data frame.

```{r}
starwars <- starwars
```

- Drop all columns after `species`.

```{r}
starwars <- starwars %>% 
  select(!films:starships)
```

- Create a new column `height_m`, listing the character’s height in meters. Drop `height` afterwards.

```{r}
starwars <- starwars %>% 
  mutate(height_m = height/100) %>% 
  select(-height)
```

---

- Keep the following variables: `name`, `mass`, all `_color` variables as well as all numeric variables.

```{r}
starwars <- starwars %>% 
  select(name, mass, 
         contains("_color"),
         where(is.numeric))
```

- Set the height variable as the first column in your dataset. Put all the `_color`-variables at the end of the dataset. Try to find a neater way to do that than just manually typing all the `_color`-variables. Tip: Check out the `relocate()`-helpfile.

```{r}
starwars <- starwars %>% 
  relocate(height_m) %>% 
  relocate(contains("_color"), .after = last_col())
```

---

- Bonus question: Which characters were born more than 100 years before the Battle of Yavin?

```{r, results='markup'}
starwars %>% 
  filter(birth_year > 100) %>% 
  pull(name)
```

```{r, include=FALSE}
pagedown::chrome_print("dplyr1_solutions.html")
```



