---
title: "Tidyverse - Solutions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE, warning = FALSE, results = "hide") #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
library(tidyverse)
library(palmerpenguins)
data(package = "palmerpenguins")
library(knitr)
library(kableExtra)
```

## Exercises

- Install the Tidyverse or update it. Then load it.

```{r}
#install.packages("tidyverse") #if not installed/updated yet
library(tidyverse)
```

- Load the `magrittr` package and check out its help file on the forward pipe operator (`%>%`).

```{r}
library(magrittr)
?magrittr
```

---

- Let’s compare the Base R approach with the Tidyverse approach:

- First, install the package `palmerpenguins` and load the built-in penguins dataset into RStudio. You can do this by typing `data(package = "palmerpenguins")`. How many rows and variables does the `penguins` dataset have? And how many factors?
  
```{r}
#install.packages("palmerpenguins") #if not installed yet

#load package
library(palmerpenguins)

#load the datasets
data(package = "palmerpenguins")

#dimensions of the penguins dataset
glimpse(penguins)
# --> 344 rows, 8 variables, 3 factors
```

---

- Now select the following variables: `species`, `island`, `flipper_length_mm`, `sex`, `year`, once with the Base R approach, and once with the Tidyverse approach
  
```{r}
#base r
penguins[, c("species", "island", "flipper_length_mm", "sex", "year")]

#tidyverse
penguins %>% 
select(species, island, flipper_length_mm, sex, year)
```

- Remove all NAs. Again, once with the Base R and once with the Tidyverse approach.
    
```{r}
#base r
na.omit(penguins)

#tidyverse
penguins %>% 
  drop_na()
```

---

- What’s the median flipper length of the different species? Try to obtain the same results with both approaches.

```{r, results='markup'}
#base r
penguins <- penguins[!(is.na(penguins$bill_length_mm)),]

tapply(penguins$bill_length_mm, penguins$species, median)

#tidyverse
penguins %>% 
  group_by(species) %>% 
  summarise(median_flipper_length = median(bill_length_mm, na.rm = T))
```

---

- What’s the minimum body mass on the different islands? Again, try to find the same results with both approaches.

```{r, results='markup'}
#base r
penguins <- penguins[!(is.na(penguins$body_mass_g)),]

tapply(penguins$body_mass_g, penguins$island, min)

#tidyverse
penguins %>% 
  group_by(island) %>% 
  summarise(min_bodymass = min(body_mass_g, na.rm = T))
```

```{r, include=FALSE}
pagedown::chrome_print("tidyverse_solutions.html")
```
