---
title: "Tidyverse - Exercises"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE) #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
library(tidyverse)
library(palmerpenguins)
data(package = "palmerpenguins")
library(knitr)
library(kableExtra)
```

## Exercises

- Install the Tidyverse or update it. Then load it.

- Check out magrittr’s help file on the forward pipe operator (`%>%`).

- Let’s compare the Base R approach with the Tidyverse approach:

  a. First, install the package `palmerpenguins` and load the built-in penguins dataset into RStudio. You can do this by typing `data(package = "palmerpenguins")`. How many rows and variables does the `penguins` dataset have? And how many factors?

  b. Now select the following variables: `species`, `island`, `flipper_length_mm`, `sex`, `year`, once with the Base R approach, and once with the Tidyverse approach.

  c. Remove all NAs. Again, once with the Base R and once with the Tidyverse approach.
    
---

- What’s the median flipper length of the different species? Try to obtain the same results with both approaches.

- What’s the minimum body mass on the different islands? Again, try to find the same results with both approaches.

```{r, include=FALSE}
pagedown::chrome_print("tidyverse_exercises.html")
```
