---
title: "Presentation Template"
author: 
  - "Author"
date: "Date"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, message = FALSE) #rendering sharp plots / not showing messages
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
library(tidyverse)
```

# A slide with bulletpoints

- Bulletpoint 1
- Bulletpoint 2
- Bulletpoint 3

---
# A slide with code

In order to higlight code, you have to put `#<<` after the corresponding line. Note that with the `theme_xaringan()` option, the plot design will be automatically adapted to the current presentation template!

```{r, message=FALSE, fig.align='center', fig.height=3, fig.width=3}
ggplot(mtcars) +
  aes(mpg, disp) +
  geom_point() +   #<<
  geom_smooth() +    #<<
  theme_xaringan()
```

---
#Code next to plot

If you like you can also have your plot next to your code.

.pull-left[
```{r car-plot, fig.show = 'hide'}
ggplot(mtcars) +
  aes(mpg, disp) +
  geom_point() +   #<<
  geom_smooth() +    #<<
  theme_xaringan()
```
]
.pull-right[
```{r ref.label = 'car-plot', echo = FALSE}
```
]


---
#How to show a table

```{r}
head(mtcars) %>% 
  knitr::kable(format = "html")
```

---
#Some links to customize your presentations

- Manual by the author: https://bookdown.org/yihui/rmarkdown/xaringan.html
- A lot of useful tips: https://arm.rbind.io/slides/xaringan.html#1
